---
title: "Supplementary Materials: Full Model Output"
subtitle: "Playtime–wellbeing associations are practically null across video game genres"
date: last-modified
execute:
  warning: false
  message: false
  echo: false
format:
  html:
    toc: true
    theme: cosmo
    embed-resources: true
---

This supplement reports the full posterior summaries for the four Bayesian
multilevel models in the main manuscript: M1 (total playtime REWB), M2
(pre-registered independent-coefficient genre model), M3 (hierarchical prior genre
model), and M4 (multi-membership genre model). Models are loaded from the
cached fits in `models/`; to regenerate them, render `manuscript.qmd`.

```{r}
#| label: setup
#| include: false

library(here)
library(dplyr)
library(tibble)
library(glue)
library(brms)
library(posterior)
library(tinytable)

h1_model      <- readRDS(here("models/h1_rewb_model.rds"))
h2_model      <- readRDS(here("models/h2_genre_model.rds"))
h2_hier_model <- readRDS(here("models/h2_hier_model_v2.rds"))
h2_mm_model   <- readRDS(here("models/h2_mm_model_full.rds"))

# Convert summary(brmsfit) into a single tidy table with grouped sections
# (population-level effects, group-level effects per grouping factor, and
# family-specific parameters), rendered with tinytable. Custom stanvars
# parameters (e.g. M3's tau hyperparameters) are absent from summary() and
# can be appended via extra_pars, summarised directly from the draws.
brms_summary_tt <- function(model, extra_pars = NULL) {
  s <- summary(model)

  tidy_section <- function(df, section) {
    df |>
      tibble::rownames_to_column("Parameter") |>
      as_tibble() |>
      mutate(section = section)
  }

  sections <- list(tidy_section(s$fixed, "Population-level effects"))
  for (g in names(s$random)) {
    sections <- c(sections, list(
      tidy_section(s$random[[g]], glue("Group-level effects: {g}"))
    ))
  }
  sections <- c(sections, list(
    tidy_section(s$spec_pars, "Family-specific parameters")
  ))

  if (!is.null(extra_pars)) {
    extra <- posterior::summarise_draws(
      posterior::subset_draws(posterior::as_draws_array(model), variable = extra_pars),
      Estimate = mean, Est.Error = sd,
      ~posterior::quantile2(.x, probs = c(0.025, 0.975)),
      Rhat = posterior::rhat,
      Bulk_ESS = posterior::ess_bulk, Tail_ESS = posterior::ess_tail
    ) |>
      rename(Parameter = variable, `l-95% CI` = q2.5, `u-95% CI` = q97.5) |>
      mutate(section = "Model-specific hyperparameters")
    sections <- c(sections, list(extra))
  }

  combined <- bind_rows(sections) |>
    transmute(
      section,
      Parameter,
      Estimate     = sprintf("%.3f", Estimate),
      `Est. error` = sprintf("%.3f", Est.Error),
      `95% CrI`    = sprintf("[%.3f, %.3f]", `l-95% CI`, `u-95% CI`),
      `R-hat`      = sprintf("%.3f", Rhat),
      `Bulk ESS`   = sprintf("%d", round(Bulk_ESS)),
      `Tail ESS`   = sprintf("%d", round(Tail_ESS))
    )

  # Row indices where each section starts (for tinytable group headers)
  group_starts <- combined |>
    mutate(row = row_number()) |>
    group_by(section) |>
    summarise(start = min(row), .groups = "drop") |>
    arrange(start)
  group_i <- as.list(group_starts$start)
  names(group_i) <- group_starts$section

  combined |>
    select(-section) |>
    tt() |>
    group_tt(i = group_i) |>
    format_tt(escape = TRUE) |>
    style_tt(i = 0, bold = TRUE)
}
```

# M1: Total Playtime (H1)

```{r}
#| label: tbl-supp-m1
#| tbl-cap: "Full posterior summary for M1 (total playtime REWB model). Estimates are posterior means with 95% credible intervals; R-hat and effective sample sizes (ESS) are convergence diagnostics."

brms_summary_tt(h1_model)
```

# M2: Pre-Registered Independent Genre Coefficients (H2)

```{r}
#| label: tbl-supp-m2
#| tbl-cap: "Full posterior summary for M2 (pre-registered independent-coefficient genre model). Genre parameters are labelled `gh_{genre}_within` and `gh_{genre}_between` for the within- and between-person exposure terms."

brms_summary_tt(h2_model)
```

# M3: Hierarchical Prior (H2)

```{r}
#| label: tbl-supp-m3
#| tbl-cap: "Full posterior summary for M3 (hierarchical prior genre model). `tau_genre_within` and `tau_genre_between` are the learned hyperparameters governing partial pooling of the genre coefficients."

brms_summary_tt(h2_hier_model,
                extra_pars = c("tau_genre_within", "tau_genre_between"))
```

# M4: Multi-Membership (H2)

```{r}
#| label: tbl-supp-m4
#| tbl-cap: "Full posterior summary for M4 (multi-membership genre model). Group-level standard deviations for the `mm()` term are the genre heterogeneity parameters (σ~s~ within and between)."

brms_summary_tt(h2_mm_model)
```
