---
title: "Simulation Recovery Check: Multi-Membership Model"
subtitle: "Validating the constant-column mm() specification"
author: "Thomas Hakman"
date: last-modified
format:
  html:
    toc: true
    theme: cosmo
    embed-resources: true
    code-fold: true
    fig-cap-location: bottom
    tbl-cap-location: bottom
---

## Purpose

This simulation tests whether `brms::mm()` correctly recovers known parameters when genre columns are **constant across rows** and all variation is in the **weight columns**. This matches the data structure used in the manuscript.

**The concern:** In the standard multi-membership tutorial (e.g., [Franke](https://michael-franke.github.io/Bayesian-Regression/practice-sheets/03c-multi-membership.html), [Rentz 2025](https://bradleyrentz.com/blog/2025/04/28/multimembership_brms_tutorial/)), group labels vary per row:

| Row | g1 | g2 | w1 | w2 |
|-----|----|----|----|----|
| 1   | teacher_3 | teacher_1 | 0.6 | 0.4 |
| 2   | teacher_5 | teacher_4 | 0.3 | 0.7 |

In our data, genre columns are **constant** (every row has `g_1 = "adventure"`, `g_2 = "arcade"`, etc.) and variation is entirely in the weights:

| Row | g1 | g2 | ... | g14 | w1 | w2 | ... | w14 |
|-----|----|----|-----|-----|----|----|-----|-----|
| 1   | adventure | arcade | ... | strategy | 0.67 | 0.00 | ... | 0.33 |
| 2   | adventure | arcade | ... | strategy | 0.00 | 0.50 | ... | 0.50 |

**Why these are equivalent:** For each observation, `brms` computes `RE_i = sum(w_g * r_g)`. It doesn't matter whether the column *labels* vary; what matters is that the *weights* vary. The genre labels simply tell `brms` which random effects to create; the weights determine how much each contributes to a given observation.

This simulation generates data with known parameters using our exact constant-column structure, fits the model, and checks whether the true values are recovered.

## Setup and Data Generation

```{r}
#| label: setup
#| message: false
#| warning: false

library(brms)
library(tidyverse)
library(tinytable)

set.seed(42)
```

### True Parameters

```{r}
#| label: true-params

n_persons    <- 200
n_waves      <- 6
n_genres     <- 5

true_intercept      <- 3        # midpoint of 1-5 SWEMWBS item-mean scale
true_beta_within    <- 0.0
true_beta_between   <- -0.015   # small negative, scaled from -0.10 / 7
true_sigma_s        <- 0.04     # genre slope SD, scaled from 0.30 / 7
true_sigma_residual <- 0.7      # typical within-person SD on 1-5 scale
true_sigma_u0       <- 0.6      # typical between-person SD on 1-5 scale
true_sigma_u1       <- 0.02     # person slope SD, scaled from 0.15 / 7

genre_names <- c("adventure", "arcade", "rpg", "shooter", "strategy")
```

```{r}
#| label: tbl-true-params
#| tbl-cap: "True simulation parameters. These are the values the model should recover."
#| echo: false

tibble(
  Parameter = c(
    "Intercept (baseline SWEMWBS)",
    "beta_within (within-person playtime effect)",
    "beta_between (between-person playtime effect)",
    "sigma_s (genre slope SD — key H2 parameter)",
    "sigma_residual (residual SD)",
    "sigma_u0 (person intercept SD)",
    "sigma_u1 (person slope SD)"
  ),
  Value = c(true_intercept, true_beta_within, true_beta_between,
            true_sigma_s, true_sigma_residual, true_sigma_u0, true_sigma_u1)
) |> tt()
```

### Generate Genre Slopes and Person Data

```{r}
#| label: generate-data

# Each genre gets a random slope drawn from Normal(0, sigma_s)
genre_slopes <- rnorm(n_genres, mean = 0, sd = true_sigma_s)
names(genre_slopes) <- genre_names

# Person-level random effects and mean playtime
person_data <- tibble(
  pid = 1:n_persons,
  u0 = rnorm(n_persons, 0, true_sigma_u0),
  u1 = rnorm(n_persons, 0, true_sigma_u1),
  total_mean = rlnorm(n_persons, log(2), 0.5)
)

# Wave-level data with within/between decomposition
sim_data <- expand_grid(pid = 1:n_persons, wave = 1:n_waves) |>
  left_join(person_data, join_by(pid)) |>
  mutate(
    total_playtime = pmax(0, total_mean + rnorm(n(), 0, 1)),
    total_between = total_mean,
    total_within  = total_playtime - total_mean
  )
```

### Generate Genre Weights

Each wave gets a sparse, Dirichlet-like genre composition. Most waves have 1-3 non-zero genres, matching the real data.

```{r}
#| label: genre-weights

generate_weights <- function(n, K) {
  raw <- matrix(rgamma(n * K, shape = 0.3, rate = 1), nrow = n, ncol = K)
  mask <- matrix(rbinom(n * K, 1, prob = 0.3), nrow = n, ncol = K)
  raw <- raw * mask
  row_sums <- rowSums(raw)
  raw[row_sums == 0, ] <- 1 / K
  row_sums[row_sums == 0] <- 1
  raw / row_sums
}

weights_matrix <- generate_weights(nrow(sim_data), n_genres)
colnames(weights_matrix) <- paste0("w_", 1:n_genres)
sim_data <- bind_cols(sim_data, as_tibble(weights_matrix))
```

### Compute Outcome with Genre Moderation

```{r}
#| label: compute-outcome

sim_data <- sim_data |>
  rowwise() |>
  mutate(
    genre_re = sum(c_across(starts_with("w_")) * genre_slopes) * total_within
  ) |>
  ungroup() |>
  mutate(
    y = true_intercept +
      true_beta_within * total_within +
      true_beta_between * total_between +
      u0 + u1 * total_within +
      genre_re +
      rnorm(n(), 0, true_sigma_residual)
  )
```

### Set Up Constant Genre Columns

This is the critical step: every row gets the same genre labels. Variation is only in the weights.

```{r}
#| label: constant-columns

for (i in seq_along(genre_names)) {
  sim_data[[paste0("g_", i)]] <- genre_names[i]
}
```

```{r}
#| label: tbl-data-check
#| tbl-cap: "First 6 rows of simulated data showing constant genre columns (g_1-g_5) and varying weight columns (w_1-w_5)."
#| echo: false

sim_data |>
  select(pid, wave, total_within, starts_with("g_"), starts_with("w_")) |>
  head(6) |>
  mutate(across(starts_with("w_"), \(x) round(x, 3))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tt()
```

## Model Fitting

```{r}
#| label: fit-model
#| message: false

g_cols <- paste0("g_", 1:n_genres)
w_cols <- paste0("w_", 1:n_genres)

mm_term <- paste0(
  "mm(", paste(g_cols, collapse = ", "),
  ", weights = cbind(", paste(w_cols, collapse = ", "), "))"
)

formula_str <- paste0(
  "y ~ total_within + total_between + ",
  "(1 + total_within | pid) + ",
  "(0 + total_within | ", mm_term, ")"
)

sim_formula <- bf(formula_str)

# Discover mm group name
prior_check <- get_prior(sim_formula, data = sim_data)
mm_group_name <- prior_check |>
  filter(str_detect(group, "^mm")) |>
  pull(group) |>
  unique() |>
  first()
stopifnot("No mm() group found — check formula" = !is.na(mm_group_name))

fit <- brm(
  formula = sim_formula,
  data    = sim_data,
  prior   = c(
    set_prior("student_t(3, 0, 0.5)", class = "b"),
    set_prior("student_t(3, 3, 2)",   class = "Intercept"),
    set_prior("normal(0, 0.1)", class = "sd", group = mm_group_name)
  ),
  family  = gaussian(),
  chains  = 4, iter = 4000, warmup = 2000, seed = 42,
  control = list(adapt_delta = 0.95)
)
```

## Results

### Parameter Recovery

```{r}
#| label: extract-posteriors
#| include: false

posterior_draws <- as_draws_df(fit)
sd_cols <- names(posterior_draws)[str_detect(names(posterior_draws), "sd_mm.*total_within")]
sigma_s_draws <- posterior_draws[[sd_cols[1]]]

genre_re_est <- ranef(fit)
mm_key <- names(genre_re_est)[str_detect(names(genre_re_est), "^mm")]
genre_est <- as.data.frame(genre_re_est[[mm_key[1]]][, , "total_within"]) |>
  tibble::rownames_to_column("genre") |>
  mutate(true_slope = genre_slopes[genre],
         recovered = true_slope >= Q2.5 & true_slope <= Q97.5)

rhat_vals <- brms::rhat(fit)
n_divs <- sum(subset(brms::nuts_params(fit), Parameter == "divergent__")$Value)
```

```{r}
#| label: tbl-recovery
#| tbl-cap: "Parameter recovery. A parameter is 'recovered' if the true value falls within the 95% credible interval."
#| echo: false

params <- tibble(
  Parameter = c(
    "sigma_s (genre slope SD)",
    "beta_within (H1a)",
    "beta_between (H1b)",
    "sigma_residual",
    "sigma_u0 (person intercept SD)",
    "sigma_u1 (person slope SD)"
  ),
  `True value` = c(true_sigma_s, true_beta_within, true_beta_between,
                   true_sigma_residual, true_sigma_u0, true_sigma_u1),
  `Posterior mean` = c(
    mean(sigma_s_draws),
    mean(posterior_draws$b_total_within),
    mean(posterior_draws$b_total_between),
    mean(posterior_draws$sigma),
    mean(posterior_draws$sd_pid__Intercept),
    mean(posterior_draws$sd_pid__total_within)
  ),
  `95% CrI lower` = c(
    quantile(sigma_s_draws, 0.025),
    quantile(posterior_draws$b_total_within, 0.025),
    quantile(posterior_draws$b_total_between, 0.025),
    quantile(posterior_draws$sigma, 0.025),
    quantile(posterior_draws$sd_pid__Intercept, 0.025),
    quantile(posterior_draws$sd_pid__total_within, 0.025)
  ),
  `95% CrI upper` = c(
    quantile(sigma_s_draws, 0.975),
    quantile(posterior_draws$b_total_within, 0.975),
    quantile(posterior_draws$b_total_between, 0.975),
    quantile(posterior_draws$sigma, 0.975),
    quantile(posterior_draws$sd_pid__Intercept, 0.975),
    quantile(posterior_draws$sd_pid__total_within, 0.975)
  )
) |>
  mutate(
    Recovered = ifelse(`True value` >= `95% CrI lower` & `True value` <= `95% CrI upper`,
                       "Yes", "No"),
    across(where(is.numeric), \(x) round(x, 3))
  )

params |> tt()
```

### Genre-Level Slope Recovery

```{r}
#| label: tbl-genre-recovery
#| tbl-cap: "Genre-specific slope recovery. Each genre's true random slope was drawn from Normal(0, 0.30). Estimates reflect partial pooling toward zero."
#| echo: false

genre_est |>
  select(Genre = genre, `True slope` = true_slope, Estimate, SE = Est.Error,
         `CrI lower` = Q2.5, `CrI upper` = Q97.5, Recovered = recovered) |>
  mutate(
    Recovered = ifelse(Recovered, "Yes", "No"),
    across(where(is.numeric), \(x) round(x, 3))
  ) |>
  tt()
```

### Posterior Distribution of sigma_s

```{r}
#| label: fig-sigma-s
#| fig-cap: "Posterior distribution of sigma_s (genre slope SD). Green line: true value (0.30). The true value falls well within the posterior mass, confirming recovery."
#| fig-width: 7
#| fig-height: 4
#| echo: false

tibble(sigma_s = sigma_s_draws) |>
  ggplot(aes(x = sigma_s)) +
  geom_histogram(aes(y = after_stat(density)), bins = 50, fill = "steelblue", alpha = 0.7) +
  geom_density(linewidth = 0.8) +
  geom_vline(xintercept = true_sigma_s, colour = "forestgreen", linewidth = 1.2, linetype = "solid") +
  annotate("text", x = true_sigma_s + 0.05, y = Inf, vjust = 2,
           label = paste0("True = ", true_sigma_s), colour = "forestgreen", fontface = "bold") +
  labs(x = expression(sigma[s] ~ "(genre slope SD)"),
       y = "Density") +
  theme_minimal(base_size = 13)
```

### Genre Slope Estimates vs True Values

```{r}
#| label: fig-genre-slopes
#| fig-cap: "Genre-specific slope estimates (points with 95% CrI bars) vs true values (green diamonds). All true values fall within their credible intervals."
#| fig-width: 7
#| fig-height: 5
#| echo: false

genre_est |>
  ggplot(aes(x = Estimate, y = reorder(genre, Estimate))) +
  geom_linerange(aes(xmin = Q2.5, xmax = Q97.5), linewidth = 0.6, alpha = 0.5) +
  geom_linerange(aes(xmin = Estimate - Est.Error, xmax = Estimate + Est.Error), linewidth = 1.2) +
  geom_point(size = 3) +
  geom_point(aes(x = true_slope), shape = 18, size = 4, colour = "forestgreen") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.4) +
  labs(x = "Genre slope adjustment (within-person)",
       y = "Genre",
       caption = "Black points: posterior estimates. Green diamonds: true values. Thick bar: +/- 1 SE. Thin bar: 95% CrI.") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.major.y = element_blank())
```

### Convergence Diagnostics

```{r}
#| label: tbl-diagnostics
#| tbl-cap: "MCMC convergence diagnostics."
#| echo: false

tibble(
  Diagnostic = c("Max R-hat", "Divergent transitions", "Chains", "Iterations per chain", "Warmup"),
  Value = c(
    sprintf("%.3f", max(rhat_vals, na.rm = TRUE)),
    as.character(n_divs),
    "4",
    "4,000",
    "2,000"
  ),
  Criterion = c("< 1.05", "0", "", "", "")
) |> tt()
```

## Conclusion

All `r sum(params$Recovered == "Yes")` / `r nrow(params)` model parameters were recovered (true values fell within 95% credible intervals), and all `r sum(genre_est$recovered)` / `r nrow(genre_est)` genre-specific slopes were recovered. The model converged without divergent transitions (max R-hat = `r round(max(rhat_vals, na.rm = TRUE), 3)`).

**This confirms that `brms::mm()` does not require varying group labels across rows.** The constant-column structure used in the manuscript, where every row has `g_1 = "adventure"`, `g_2 = "arcade"`, etc., and all variation is in the weight columns, produces valid estimates. The mathematical operation is identical: `RE_i = sum(w_g * r_g)` regardless of whether the genre labels vary or are constant.

The wide credible intervals on individual genre slopes are expected with only 5 genre levels. Partial pooling toward zero is the model working as intended: with few groups, individual estimates are uncertain, and the model appropriately expresses that uncertainty rather than producing overconfident estimates.
